BETTING AND BELIEF: PREDICTION MARKETS AND ATTRIBUTION OF CLIMATE CHANGE

### 2016-06-29 prediction_market_wsc.Rnw
Some paper edits.

\citeN{Bloch_2010} have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change. Existing prediction markets (e.g. \href{https://hypermind.com/hypermind/app.html}{hypermind}, \href{https://www.betfair.com/exchange/}{betfair}, and \href{https://www.predictit.org/}{PredictIt}) focus on near term events such as elections months away, so it is difficult to extrapolate empirical findings to the climate case. Furthermore, we are interested in investigating the unobservable beliefs of traders. Therefore, we turn to simulation modeling informed by climate and economic theory. We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what social and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs. Our work can be extended as part of a computational design process for effective climate prediction markets.

@@

In our model, traders bet on global temperature anomalies six years in the future.
During the six-year period, traders buy and sell futures.
Every year, traders update their models and forecast of future temperatures based on newly available data.
At the end of each six-year period, winners collect gains and traders revise beliefs about climate models, based on their ideology and the beliefs of top earners in their social network.
In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market procedures, and the social network connecting agents.
We conclude with details on model dynamics and an overview of the model parameters we experimentally vary. The model code and a full \href{https://github.com/JohnNay/predMarket/blob/master/odd.pdf}{Overview, Design concepts and Details (ODD) specification} can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.

@@

For climate time-series, we use the annual anomaly of global mean temperature. For years from 1880--2014 we use the GISTEMP global mean land-sea annual temperature anomalies \shortcite{GISTEMP_2016,Hansen_2010} and for years from 2015 onward, we project future climates under two alternative theories: In both theories, changes in global temperature are proportional to changes in a deterministic forcing plus a stochastic noise term. For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million \shortcite[p.~37]{Archer_2012}, and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle \shortcite{Soon_2005}.  Most scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more. In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest single forcing \shortcite[p.~14]{IPCC_WG1_AR5_2013}, so for simplicity we consider only \ce{CO2}. Choosing only one forcing term for each of the competing models simplifies comparison because each model has the same number of adjustable parameters.

>>>>

\citeN{Bloch_2010} have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change. Existing prediction markets (e.g. \href{https://hypermind.com/hypermind/app.html}{hypermind}, \href{https://www.betfair.com/exchange/}{betfair}, and \href{https://www.predictit.org/}{PredictIt}) focus on near term events such as elections months away, so it is difficult to extrapolate empirical findings to the climate case. Furthermore, we are interested in investigating the unobservable beliefs of traders. Therefore, we turn to simulation modeling informed by climate and economic theory. We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what social and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs. Our work can be extended as part of a computational design process for effective climate prediction markets and we release all our code along with this paper.

@@

In our model, traders bet on global temperature anomalies six years in the future.
During the six-year period, traders buy and sell futures.
Every year, traders update their models and forecast of future temperatures based on newly available data.
At the end of each six-year period, winners collect gains and traders revise beliefs about climate models, based on their ideology and the beliefs of top earners in their social network.
In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market procedures, and the social network connecting agents.
We conclude with details on model dynamics and an overview of the model parameters we experimentally vary. A full Overview, Design concepts and Details (ODD) specification can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.

@@

For climate time-series, we use the annual anomaly of global mean temperature. For years from 1880--2014 we use the GISTEMP global mean land-sea annual temperature anomalies \shortcite{GISTEMP_2016,Hansen_2010} and for years from 2015 onward, we project future climates under two alternative theories: In both theories, changes in global temperature are proportional to changes in a deterministic forcing plus a stochastic noise term. For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million \shortcite[p.~37]{Archer_2012}, and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle \shortcite{Soon_2005}.  Most scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more. In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest single forcing \shortcite[p.~14]{IPCC_WG1_AR5_2013}. Choosing only one forcing term for each of the competing models simplifies comparison because each model has the same number of adjustable parameters, therefore we consider only \ce{CO2}.
